Measuring sustainability with Australian vine growers using a Bayesian Network

### 2023-10-31 article.tex
Notes for Korb & Nicholson about KEBN

Korb & Nicholson describe three major issues in the creation of Bayesian Networks though knowledge Engineering:
- Generating Arc Structure
- Parametarisation
- dealing with utilities

@@

\subsection{The problem}
\subsubsection{Not understanding the problem context, complexity without value.}
It is crucial to gain a clear understanding of the problem and its context. This can be seen as answering the following questions:
\begin{enumerate}
        \item what do you want to reason about?
        \item What dont you know?
        \item What infomration do you have?
        \item What do you know?
These questions are well addressed in prior research. The creation of baseline models showed strong linear relationships between many variables. The creation of XGBoosted trees uncovered the ability to well inform any particular variable but with no real causal structure or understandability. here we seek to pursue causality and link these variables together in a way that is more aligned with a decision support structure than purely as a predictive exercise.
\end{enumerate}
\subsection{Structure - Nodes}
\subsubsection{Getting the node values wrong. An exhaustive list below}
\begin{enumerate}
        \item Node values aren't exhaustive
        \item Node values aren't exhaustive
        \item Node values aren't mutually exclusive
        \item Incorrect modelling of mutually exclusive outcomes
        \item Trying to model fuzzy categories
        \item Confusing state and Probability
        \item Confusion about what the node represents
\end{enumerate}
Two major strategies were employed to overcome these problems. Tying nodes explicitly to a data source that corroberated the definition and state in some manner as the node. For node s that did not have data set they were sought to be defined with an explicit explanation that experts agreed upon or had some external explanation such as legislation.
\subsection{Structure Arcs}
\subsubsection{Getting the arc directions wrong}
Arc directions are commonly wrong for the following reasons:
\begin{enumerate}
        \item Modelling reasoning rather than causation
        \item Inverting cause and effect
        \item Missing variables
\end{enumerate}
To note: Korb and Nicholson do not actually suggest a method for overcoming these problems. There is an inherint risk in creating a model and not having knowledge of a missing variable or of a causation. It is important to acknowledge that there is always a possibility that causation is mixed up with reasoning, and that not all variables are always known.
\subsubsection{Too many parents}
Sometimes it is unavoidable that a node is affected by many factors, or other nodes. There are some methods to help overcome this such as divorcing, in which a node is created to capture the cumulative affects of several parents on a child. This technique was employed in the creation of the structure of the BN with the expert panel.
\subsection{Parameters}
\subsubsection{Experts estimates of probabilities are biased}
-       -       - overconfidence
-       -       - Anchoring
-       -       - Availability
-       -       - Inconsistent "filling in" of large CPTs
\subsubsection{Incoherent probabilities (not summing to 1)}
-       -       - Being dead certain.

-

Boneh 2010 describe Knowledge Engineering with Bayesian Networks (KEBN) as occurring in a 4 stage life cycle:
- Bayesian Network Structure
- Probability Parameters
- Decision Structure
- Utilities Preference

>>>>

\citep{korbBayesianArtificialIntelligence2011} Korb & Nicholson describe three major issues in the creation of Bayesian Networks though knowledge Engineering:
- Generating Arc Structure
- Parametarisation
- dealing with utilities

@@

\subsection{The problem}
\subsubsection{Not understanding the problem context, complexity without value.}
It is crucial to gain a clear understanding of the problem and its context. This can be seen as answering the following questions:
\begin{enumerate}
        \item what do you want to reason about?
        \item What dont you know?
        \item What infomration do you have?
        \item What do you know?
These questions are well addressed in prior research. The creation of baseline models showed strong linear relationships between many variables. The creation of XGBoosted trees uncovered the ability to well inform any particular variable but with no real causal structure or understandability. here we seek to pursue causality and link these variables together in a way that is more aligned with a decision support structure than purely as a predictive exercise.
\end{enumerate}
\subsection{Structure - Nodes}
\subsubsection{Getting the node values wrong. An exhaustive list below}
\begin{enumerate}
        \item Node values aren't exhaustive
        \item Node values aren't exhaustive
        \item Node values aren't mutually exclusive
        \item Incorrect modelling of mutually exclusive outcomes
        \item Trying to model fuzzy categories
        \item Confusing state and Probability
        \item Confusion about what the node represents
\end{enumerate}
Two major strategies were employed to overcome these problems. Tying nodes explicitly to a data source that corroberated the definition and state in some manner as the node. For node s that did not have data set they were sought to be defined with an explicit explanation that experts agreed upon or had some external explanation such as legislation.
\subsection{Structure Arcs}
\subsubsection{Getting the arc directions wrong}
Arc directions are commonly wrong for the following reasons:
\begin{enumerate}
        \item Modelling reasoning rather than causation
        \item Inverting cause and effect
        \item Missing variables
\end{enumerate}
To note: Korb and Nicholson do not actually suggest a method for overcoming these problems. There is an inherint risk in creating a model and not having knowledge of a missing variable or of a causation. It is important to acknowledge that there is always a possibility that causation is mixed up with reasoning, and that not all variables are always known.
\subsubsection{Too many parents}
Sometimes it is unavoidable that a node is affected by many factors, or other nodes. There are some methods to help overcome this such as divorcing, in which a node is created to capture the cumulative affects of several parents on a child. This technique was employed in the creation of the structure of the BN with the expert panel.
\subsection{Parameters}
\subsubsection{Experts estimates of probabilities are biased}
\begin{enumerate}
        \item Overconfidence: The attribution of higher than justifiable probabilities to nodes/events.
        \indent The tendency to attribute higher than justifiable probabilities (\citep{lichtensteinCalibrationProbabilitiesState1982} 
        \item Anchoring: The tendency towards higher/lower based off of an arbitrary point of comparison. "Anchoring" people at a higher or lower than normal prediction because of some assumption or example  \citep{kahnemanPsychologyPrediction1973}.
        \item Availability: The assessing of an event as more probable than is justifiable becaues it is more memorable or noticeable \citep{tverskyJudgmentUncertaintyHeuristics1974}.

\end{enumerate}
We have attempted to overcome this with two methods. A diverse and varried expert panel. Those participanting have experience in multiple vineyards as managers and other duties as well as the panel itself having varied manager/owners from different regions and scale. With some experts being more specialised in specific areas such as agro-chemicals.
\subsubsection{Incoherent/Incoherent probabilities}
The collection of probabilities for when the CPT is large can result in inconsistent probabilities. The collection of probabilities at different times or if probabilities are supplied by different experts can result in inconsistent probabilities as well.
The use of desciptors such as very high or low can be of assistance \cite{vandergaag2013elicit}. In this study we overcame this through the use of worded response such as high and low. We then assigned probabilities to those broader cases. We had every expert agree upon each of the probabilities so that they could argue for or against any particular bias and so that the probabilities would even out to a more broader applicability, across all of the experts experiences.
% \item Inconsistent "filling in" of large CPTs:
-       -       - Being dead certain.

\citep{morganUncertaintyGuideDealing1990}
Two attributes of good elicitation process are:
\begin{enumerate}
        \item The expert should be apprised of what is known about the process, especially the nearly universal tendency to overcomefidence and other forms of bias. In order to avoid some problems, values should be elicited in random order and the expert not given feedback on how the different values fit together until a complet set has been elicited.
        \item The elicitation process is not simply one of requesting and recording numbers but also one of the refining the definitions of variables and terms to be used in the model. What values are elicited depends directly upon the interpretation of terms and these should be made as explicit as possible and recorded during the elicitation. This is part of the process management described earlier.
\end{enumerate}
-

\citep{Boneh2021} Boneh, T. (2010). Ontology and Bayesian decision networks for supporting the meterological forecasting process. Ph. D thesis, Clayton School of infomration technology, Monash university
Boneh 2010 describe Knowledge Engineering with Bayesian Networks (KEBN) as occurring in a 4 stage life cycle:
- Bayesian Network Structure
- Probability Parameters
- Decision Structure
- Utilities Preference
